Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-09-07 paper.Rnw
knit paper after changing file names

# Apply LIME with various tuning parameters to the sine data
if (!file.exists("../data/sine_lime_explain.rds")) {

@@

    saveRDS(sine_lime_explain, "../data/sine_lime_explain.rds")

@@

    sine_lime_explain <- readRDS("../data/sine_lime_explain.rds")

@@

# Load the hamby data
bullet_train <- read.csv("../data/hamby173and252_train.csv")
bullet_test <- read.csv("../data/hamby224_test.csv")

# Extract the features and order them based on feature importance
bullet_features <- rownames(bulletxtrctr::rtrees$importance)
bullet_features_ordered <-
  data.frame(
    feature = rownames(bulletxtrctr::rtrees$importance),
    MeanDecreaseGini = bulletxtrctr::rtrees$importance
  ) %>%
  arrange(desc(MeanDecreaseGini)) %>%
  mutate(
    feature = fct_recode(feature,
      "Cross Correlation Function" = "ccf",
      "Consecutively Matching Striae" = "cms",
      "Matches" = "matches",
      "Mismatches" = "mismatches",
      "Non-Consecutively Matching Striae" = "non_cms",
      "Rough Correlation" = "rough_cor",
      "Distance Standard Deviation" = "sd_D",
      "Distance" = "D",
      "Sum of Peaks" = "sum_peaks"
    )
  )

# Create an ordered version of the data
bullet_features_ordered <- bullet_features_ordered %>%
  mutate(feature = factor(feature, levels = bullet_features_ordered$feature))

@@

# Load the signatures data
signatures <- readRDS("../data/signatures.rds")

@@

# Create or load the parallel coordinate plot
if (!file.exists("./figure-static/figure-10-1.png")) {

@@

  # Save the plot
  ggplot2::ggsave(
    plot = bullet_pcp,
    filename = "./figure-static/figure-10-1.png",
    width = 7.5,
    height = 3.5,
    units = "in",
    dpi = 1200
  )

  # Load the figure and then save as an EPS file
  bullet_pcp <- image_read("./figure-static/figure-10-1.png")
  image_write(bullet_pcp, path = "./figure-static/figure-10-1.eps", format = "eps")

@@

   # Load and print the plot
  knitr::include_graphics("./figure-static/figure-10-1.eps")

@@

# Apply LIME to all but two cases without returning the permutations
if (file.exists("../data/hamby_lime.rds")){

  # Load the files
  bullet_explain_noperms <- readRDS("../data/hamby_explain.rds")

@@

  # Save the output objects
  saveRDS(object = bullet_lime_noperms,
          file = "../data/hamby_lime.rds")
  saveRDS(object = bullet_explain_noperms,
          file = "../data/hamby_explain.rds")

>>>>

# Apply LIME with various tuning parameters to the sine data
if (!file.exists("data-sine-lime-explain.rds")) {

@@

    saveRDS(sine_lime_explain, "data-sine-lime-explain.rds")

@@

    sine_lime_explain <- readRDS("data-sine-lime-explain.rds")

@@

# Files that will need to be removed once I know how to 
#bullet_train <- read.csv("../data/hamby173and252_train.csv")
no_tank_rash = read.csv("data-no-tank-rash.csv")

# Load the bullet matching training data and make some adjustments:
#   1. filter out observations know to have tank rash
#   2. exclude features not used to train the rtrees random forest 
#   3. change the name of the same_source variable to match the test data
#   4. reorder the variables
bullet_train <-
  read.csv("data-bullet-train.csv") %>%
  slice(no_tank_rash$case) %>%
  #filter(land_id1 == "Hamby173-BrUnk-B2-L1")
  select(-lag,-abs_lag,-signature_length) %>%
  mutate(case = factor(1:n())) %>%
  rename("samesource" = "same_source") %>%
  select(case, land_id1, land_id2, everything())

# Load the bullet test data
bullet_test <- read.csv("data-bullet-test.csv")

# Extract the features
bullet_features <- rownames(bulletxtrctr::rtrees$importance)

# Add a variable to the training data with the random forest probability
# in support of a match
bullet_train$rfscore <- 
  bulletxtrctr::rtrees %>%
  predict(bullet_train %>% select(all_of(bullet_features)), type = "prob") %>%
  data.frame() %>%
  pull(TRUE.)

# Order the features based on feature importance
bullet_features_ordered <-
  data.frame(
    feature = rownames(bulletxtrctr::rtrees$importance),
    MeanDecreaseGini = bulletxtrctr::rtrees$importance
  ) %>%
  arrange(desc(MeanDecreaseGini)) %>%
  mutate(
    feature = fct_recode(feature,
      "Cross Correlation Function" = "ccf",
      "Consecutively Matching Striae" = "cms",
      "Matches" = "matches",
      "Mismatches" = "mismatches",
      "Non-Consecutively Matching Striae" = "non_cms",
      "Rough Correlation" = "rough_cor",
      "Distance Standard Deviation" = "sd_D",
      "Distance" = "D",
      "Sum of Peaks" = "sum_peaks"
    )
  )

# Create an ordered version of the data
bullet_features_ordered <- bullet_features_ordered %>%
  mutate(feature = factor(feature, levels = bullet_features_ordered$feature))
@

@@

# Load the signatures data
signatures <- read.csv("data-example-signatures.csv")

@@

# Create or load the parallel coordinate plot
if (!file.exists("figure-10-1.png")) {

@@

  # Save the plot
  ggplot2::ggsave(
    plot = bullet_pcp,
    filename = "figure-10-1.png",
    width = 7.5,
    height = 3.5,
    units = "in",
    dpi = 1200
  )

  # Load the figure and then save as an EPS file
  bullet_pcp <- image_read("figure-10-1.png")
  image_write(bullet_pcp, path = "figure-10-1.eps", format = "eps")

@@

   # Load and print the plot
  knitr::include_graphics("figure-10-1.eps")

@@

# Apply LIME to all but two cases without returning the permutations
if (!file.exists("data-bullet-lime.rds") | !file.exists("data-bullet-explain.rds")){

@@

  # Save the output objects
  saveRDS(object = bullet_lime_noperms, file = "data-bullet-lime.rds")
  saveRDS(object = bullet_explain_noperms, file = "data-bullet-explain.rds")

  # Load the files
  bullet_explain_noperms <- readRDS("data-bullet-explain.rds")
